Soundness Proofs for Iterative Deepening

### 2016-08-23 oopsla-16/proof/di.tex
prepare proof for submission and begin writing system model

>>>>

\subsection{System Model}

To reason about programs at the executable level, we will define them as a list of assembly-like instructions,
including certain concurrency primitives which we evaluate atomically, assuming the runtime provides correct implementations for them.
\begin{eqnarray*}
	\mathcal{P} &::=& [\mathcal{I}] \\
	\mathcal{I} &::=& v \leftarrow \mathsf{read}(a) \quad | \quad \mathsf{write}(a,v) \quad | \quad \mathsf{xchg}(a,v) \\
	&|& a \leftarrow \mathsf{malloc}(n) \quad | \quad \mathsf{free}(a) \\
	&|& \mathsf{mutex\_lock}(m) \quad | \quad \mathsf{mutex\_unlock}(m) \\
	&|& \mathsf{deschedule} \quad | \quad \mathsf{make\_runnable}(t) \\
	&|& v \leftarrow \mathsf{thread\_fork} \quad | \quad \mathsf{thread\_exit} \\
	&|& \mathsf{local}
\end{eqnarray*}

For the sake of brevity, we will not explicitly define execution semantics, though they are straightforward:
The execution state consists of any number of threads, each represented by a list of instructions to execute ($[\mathcal{I}]$),
a stack of value maps of thread-local variables ($v$),
and an indicator of whether the thread is runnable.
The state also includes a map of global memory, indexed by addresses denoted $a$,
and a set of mutex objects, denoted $m$.
To execute one step, arbitrarily choose one runnable thread, and evaluate the first element of its instruction list.
This gives rise to thread concurrency which can interleave at instruction granularity,
although it assumes a sequentially-consistent memory model \cite{tsopso}.
\begin{eqnarray*}
	\mathcal{E} &::=& [\mathcal{T}] \times [a \rightarrow \mathsf{int}] \times [m \rightarrow \mathsf{bool}] \\
	\mathcal{T} &::=& [\mathcal{I}] \times [v \rightarrow \mathsf{int}] \times \mathsf{bool}
\end{eqnarray*}

{\bf Memory.} $\mathsf{read}$, $\mathsf{write}$, and $\mathsf{xchg}$ access global or heap memory shared by all threads, indicated by some address $a$, and copying and/or modifying a thread-local variable $v$. % cmpxchg, xadd, etc omitted for brevity
$\mathsf{malloc}$ and $\mathsf{free}$ provide access to fresh memory accessible by all threads. % free is irrelevant to the system model in terms of expressive power

{\bf Threads.} $\mathsf{mutex\_lock}$ and
$\mathsf{mutex\_unlock}$ provide mutual exclusion:
a thread cannot evaluate $\mathsf{mutex\_lock}$ until that lock is released.
$\mathsf{deschedule}$ and $\mathsf{make\_runnable}$ allow threads to manipulate their runnability,
and $\mathsf{thread\_fork}$ and $\mathsf{thread\_exit}$ allow creation and destruction of new threads.
$\mathsf{thread\_fork}$ is defined in the Pebbles manner \cite{kspec};
we omit higher-level abstractions such as $\mathsf{create}$ or $\mathsf{join}$, which can be implemented using these primitives \cite{thrlib}.

{\bf Local state.}
$\mathsf{local}$ represents any thread-local instruction, such as flow control, function definition/calling, and modifying local variables.
For brevity, we omit a detailed list, although note that $\mathsf{call}$ would be implemented by modifying the instruction stream $[\mathcal{I}]$ and creating a new frame on the stack of local variables.

### 2016-08-23 oopsla-16/proof/proof.tex
Import proof
prepare proof for submission and begin writing system model

%
% LaTeX template for prepartion of submissions to PLDI'15
%
% Requires temporary version of sigplanconf style file provided on
% PLDI'15 web site.
%
\documentclass[pldi]{../sigplanconf-pldi15}

@@

First, we prove that given enough time, Iterative Deepening will converge to the same degree of schedule coverage that would be provided by a na\"{i}ve model checker (MC) preempting on every single instruction.
In other words, when Iterative Deepening finishes generating new data-race PPs, and completes all associated state spaces,
it serves as a full verification of {\em all possible} thread schedules of the given test case.

@@

This document is supplemental material to our main paper; we assume the reader is already familiar with our motivation and terminology.
\end{abstract}

@@

{\em removed for double blind review}
% Michael J. ``Sully'' Sullivan, Carlo Angiuli, and Nathaniel Filardo double-checked the proof and provided valuable feedback on notation style and the flow of explanation.

>>>>

%
% LaTeX template for prepartion of submissions to PLDI'15
%
% Requires temporary version of sigplanconf style file provided on
% PLDI'15 web site.
%
%\documentclass[pldi]{../sigplanconf-pldi15}
\documentclass[10pt,times,numbers]{../sigplanconf}
\usepackage[proof]{../oopsla2016}

@@

First, we prove that given enough time, Iterative Deepening will converge to the same degree of schedule coverage that would be provided by a na\"{i}ve model checker (MC) preempting on every single instruction.
In other words, when Iterative Deepening finishes generating new data-race PPs, and completes all associated state spaces,
it serves as a full verification of {\em all possible} schedules of the given test case.

@@

	This document is supplemental material to our main paper \cite{quicksand}; we assume the reader is already familiar with our motivation and terminology.
\end{abstract}

@@

%{\em removed for double blind review}
Michael J. ``Sully'' Sullivan, Carlo Angiuli, and Nathaniel Filardo double-checked the proof and provided valuable feedback on notation style and the flow of explanation.
Thanks as well to the anonymous OOPSLA reviewer \#4, whose suggestions helped improve the formality of our definitions.
This work was supported in part by the U.S. Army Research Office under grant number W911NF0910273.
